---
title: "P8105 Homework 3 [td2704]"
output: github_document
---


```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(readxl)
library(lubridate)
library(dplyr)
library(p8105.datasets)

# knitr::opts_chunk$set(
#   fig.width = 6,
#   fig.asp = .6,
#   out.width = "90%"
# )

```

```{r}

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


## Problem 1

```{r}

data("instacart")

```


```{r}

instacart

```

### Describing the instacart dataset.

The instacart dataset has `r nrow(instacart)` rows of `r length(colnames(instacart))`
variables: `r colnames(instacart)`. It is a tibble that lists products in each
order with an associated order_number and user_id, along with other order info
such as when they last ordered and what time this order was placed. It covers
a range of `r length(unique(pull(instacart,product_id)))` products in 
`r length(unique(pull(instacart,department)))` departments from 
`r length(unique(pull(instacart,order_id)))` orders by `r length(unique(pull(instacart,user_id)))` 
users. For example, the first row of the dataset describes `r pull(instacart[1,],product_name)`
from order `r pull(instacart[1,],order_id)` by user `r pull(instacart[1,],user_id)`; it is
described to be in the `r pull(instacart[1,],aisle)` aisle in the `r pull(instacart[1,],department)`
department. It was added to cart by user `r pull(instacart[1,],user_id)` `r pull(instacart[1,],add_to_cart_order)`
time(s); and it had been `r pull(instacart[1,],days_since_prior_order)` days since this user
last ordered. As another example of a row of data in this dataset, the 10th row of the dataset describes `r pull(instacart[10,],product_name)` from order `r pull(instacart[10,],order_id)` by user `r pull(instacart[10,],user_id)`; it is described to be in the `r pull(instacart[10,],aisle)` 
aisle in the `r pull(instacart[10,],department)` department. It was added to cart by user `r pull(instacart[10,],user_id)` `r pull(instacart[10,],add_to_cart_order)` time(s); 
and it had been `r pull(instacart[10,],days_since_prior_order)` days since this user
last ordered.


### How many aisles are there, and which aisles are the most items ordered from?

There are `r length(unique(pull(instacart,aisle)))` aisles in this dataset.

```{r}

aisle_info = 
  instacart %>%
  group_by(aisle, aisle_id) %>%
  count(aisle) %>%
  arrange(desc(n)) %>%
  rename(num_items_ordered = n)
  
sum(pull(aisle_info,num_items_ordered)) # to make sure we total the number of rows in dataset

aisle_info
  

```

It seems to be that the top 5 aisles most ordered from are `r pull(head(aisle_info,5),aisle)`.


### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

We can use our aisle_info dataframe from above for this.

```{r}

aisle_info %>%
  filter(num_items_ordered > 10000) %>%
  ggplot(aes(x=num_items_ordered, y=reorder(aisle, num_items_ordered))) +
  geom_bar(stat='identity') + 
  ggtitle("Number of items ordered in each aisle (more than 10000)") +
  xlab("Number items ordered") + ylab("Aisle name")

```


### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}

instacart %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  group_by(aisle, aisle_id,product_name, product_id) %>%
  count(product_name) %>%
  rename(num_items_ordered=n) %>%
  arrange(desc(num_items_ordered)) %>%
  group_by(aisle) %>%
  slice_max(num_items_ordered,n=3) %>%
  arrange(desc(num_items_ordered)) %>%
  knitr::kable()

```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}

instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  select(product_name,product_id,order_dow,order_hour_of_day) %>%
  group_by(product_name, product_id, order_dow) %>%
  summarize(mean_hour_of_day = mean(order_hour_of_day,
                                    .groups=c(product_name, product_id, order_dow))) %>%
  pivot_wider(
    names_from=order_dow,
    values_from=mean_hour_of_day) %>%
  knitr::kable(digits=2)

```


## Problem 2

```{r}
data("brfss_smart2010")

```

```{r}

brfss_smart2010

```

### Some data cleaning for BRFSS dataset.

```{r}

brfss = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic=="Overall Health") %>%
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))
  ) %>%
  rename(state = locationabbr) %>%
  rename(location = locationdesc)

# unique(pull(brfss,response)), responses already exist only from Excellent to Poor

brfss

```

### In 2002, which states were observed at 7 or more locations? What about in 2010?

2002: states observed at 7 or more locations

```{r}

brfss %>%
  filter(year == 2002) %>%
  group_by(state) %>%
  summarize(location_count = n_distinct(location)) %>%
  filter(location_count >= 7) %>%
  knitr::kable()
  

```

2010: states observed at 7 or more locations


```{r}

brfss %>%
  filter(year == 2010) %>%
  group_by(state) %>%
  summarize(location_count = n_distinct(location)) %>%
  filter(location_count >= 7) %>%
  knitr::kable()

```

### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r}

data_value_overyears_df = 
  brfss %>%
  filter(response == "Excellent") %>%
  group_by(year, state) %>%
  summarize(
    mean_data_value = mean(data_value))  %>%
  select(year, state, mean_data_value)

data_value_overyears_df
```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r, fig.height=6, out.width = "90%"}

data_value_overyears_df %>%
  ggplot(aes(x=year,y=mean_data_value)) +
  geom_line(aes(group=state,color=state))

```


### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r}

brfss %>%
  filter(state == "NY") %>%
  filter(year == 2006 | year == 2010) %>%
  ggplot(aes(x=response,y=data_value,fill=year)) + 
  geom_bar(stat='identity') + 
  facet_grid(. ~ year) +
  ggtitle("Distribution of data_value for various responses in NY in 2006 vs 2010") +
  xlab("Response") + ylab("data_value") + 
  theme(legend.position = "none")
  

```


##  Problem 3

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc)


```{r}

accel_df = 
  read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    weekend = (day == "Saturday"  | day == "Sunday")
  ) %>%
  relocate(week, day_id, day, weekend)

accel_df

```



